DOpElib:\\ Differential Equations and Optimization Environment

### 2013-12-13 Examples/Documentation/examples.tex
Update Doku

\chapter{Example Handling, Creating new Examples}
\label{chap:howtoex}
To implement new examples or to use existing examples 
from the library for own research, the user 
can simply copy an existing example. In this 
new example, own code and changes can be compiled.

Before being able to change and compile the new code, the user must 
follow some easy steps in order to modify the information related to the old code. In 
this section we explain how to modify such information using as model 
\texttt{PDE/StatPDE/Example1}.

>>>>

\chapter{Example Handling, Creating new Examples}
\label{chap:howtoex}
\section{Getting started}
Beside the fact that \dope{} is still under development,
it offers already various different (linear and nonlinear) 
examples for a lots 
of different applications in two and three dimensions; 
we refer the reader to the 
next two Chapters \ref{PDE} and \ref{OPT}.

To implement new examples or to use existing examples 
from the library for own research, the user 
can simply copy an existing example. In this 
new example, own code and changes can be compiled. Here is some 
advice to get started:
\begin{itemize}
\item If you are a first time user of \dope{} 
with some numerics background, you 
might be familiar with the Poisson (or more general Laplace) equation.
\dope{} has it too. Check-out Example \ref{PDE_Stat_Laplace_2D},
to see how \dope{} implements this well-known equation in two
dimensions or \ref{PDE_Stat_Laplace_3D} for its three-dimensional version.
\item Before you implement a new example, please check which 
existing example might be similar to your goals and get familiar 
to it. Then, proceed as described in Section \ref{getting_started}. 
\end{itemize}
\begin{remark}[Why two examples for the Laplace equation?]
\dope{} offers dimensional-independent programming! So, we could have 
implemented the 2d and 3d case in one file. The reason why we
present
to separate programs \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}
is, that in the 3d case, we show additional 
capabilities of the library itself; specifically how to use different
iterative linear solvers.
\end{remark}

\section{Creating new examples}
\label{getting_started}
Before being able to change and compile the new code, the user must 
follow some easy steps in order to modify the information related to the old code. In 
this section we explain how to modify such information using as model 
\texttt{PDE/StatPDE/Example1}.

### 2013-12-13 Examples/Documentation/foreword.tex
Update Doku

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
At the present stage the following features are supported by the library
\begin{itemize}
\item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
\item Various time stepping schemes (based on finite differences), 
  such as forward Euler, backward Euler,
  Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
\item All finite elements of from deal.II including hp-support.
\item Several examples showing the solution of several PDEs including
   Poisson, Navier-Stokes, plasticity, fluid-structure interaction problems,
and finally a coupled Biot-Lam\'e-Navier system.
\item Self-written line search and trust region newton algorithms for the 
   solution of optimization problems with PDEs \cite{NoWr00}
\item Interface to SNOPT for the solution of optimization problems with PDEs and
  additional other constraints.
\item Several examples showing how to solve various kinds of optimization problems
  involving stationary PDE constraints.
\item Mesh adaptation and goal-oriented error estimation with
the dual-weighted residual (DWR) method.
\item Different spatial triangulations for control and state variables.
\end{itemize}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
At the present stage the following features are supported by the library
\begin{itemize}
\item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
\item Various time stepping schemes (based on finite differences), 
  such as forward Euler, backward Euler,
  Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
\item All finite elements of from deal.II including hp-support.
\item Self-written line search and trust region newton algorithms for the 
   solution of optimization problems with PDEs \cite{NoWr00}
\item Interface to SNOPT for the solution of optimization problems with PDEs and
  additional other constraints.
\item Several examples showing how to solve various kinds of optimization problems
  involving stationary PDE constraints.
\item Mesh adaptation and goal-oriented error estimation with
the dual-weighted residual (DWR) method.
\item Different spatial triangulations for control and state variables.
\item Several examples showing the solution of several PDEs including
   Poisson, Navier-Stokes, plasticity, fluid-structure interaction problems,
a coupled Biot-Lam\'e-Navier system, and finally from financial mathematics
the Black-Scholes equations.
\end{itemize}

### 2013-12-13 Examples/PDE/InstatPDE/Example2/content.tex
Update Doku

In the present example, we solve a 
nonstationary fluid-structure interaction problem in 
arbitrary Lagrangian-Eulerian (ALE) coordinates. 
The mesh motion model is based on solving a biharmonic equation
\cite{Wi11} rather than a linear-elastic model. 
The underlying equations are stated in the following:
\begin{Problem}[FSI with biharmonic mesh motion]
  \label{eq:fsi:ale:biharmonic}
  Find $\{\hat v,\hat u,\hat w,\hat p\} \in \{ \hat v^D + \hat {\cal V}^0\} 
\times \{ \hat u^D + \hat {\cal V}^0 \}\times \hat {\cal V} \times \hat {\cal
  L}$, such that $\hat v (0) = \hat v^0$ and $\hat u(0) = \hat u^0$, 
for almost all time steps $t$, and
  \begin{eqnarray*}
    \begin{aligned}
      (\hat J \hat\rho_f \partial_t \hat v,\hat\psi^v)_{\hat\Omega_f}  
      +(\hat\rho_f \hat J  (\hat F^{-1}(\hat
      v-\partial_t \hat u)\cdot\hat\nabla) \hat v),
      \hat\psi^v)_{\hat\Omega_f} &\\
      + (\hat J\hat\sigma_f\hat
      F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_f}
      %
      + (\hat\rho_s \partial_t \hat v,\hat\psi^v)_{\Omega_s}  
      + (\hat J\hat\sigma_s\hat F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_s}&\\
      - \langle \hat g, \hat\psi^v \rangle_{\hat\Gamma_N} -
      (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
      &= 0&&\forall\hat\psi^v\in \hat V^0,
      \\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      (\hat\alpha_u\hat w, \hat\psi^w)_{\hat\Omega_f}  + ( \hat\alpha_u\hat\nabla\hat
      u, \hat\nabla\hat\psi^w)_{\hat\Omega_f} +( \hat\alpha_u\hat\nabla\hat
      w, \hat\nabla\hat\psi^w)_{\hat\Omega_s}  
%-\langle \hat\alpha_u \hat n_f\hat \nabla \hat u,\hat\psi^w\rangle_{\hat \Gamma_i} 
      & = 0 &&\forall\hat\psi^w\in \hat V , \\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      \hat\rho_s (\partial_t\hat u-\hat v,\hat\psi^u)_{\hat\Omega_s}
      + (\hat\alpha_u \hat \nabla \hat w,\hat \nabla\hat\psi^u)_{\hat\Omega_f}
      %-\langle \hat\alpha_w \hat n_f\hat \nabla \hat w,\hat\psi^u\rangle_{\hat \Gamma_i}
      &=0&&\forall\hat\psi^u\in \hat V^0 ,\\    
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      (\widehat{\text{div}}\,(\hat J\hat F^{-1}
      \hat v_f),\hat\psi^p)_{\hat\Omega_f} 
      + (\hat p_s ,\hat \psi^p)_{\hat\Omega_s}
      &=0&&\forall\hat\psi^p\in \hat L,
    \end{aligned}
  \end{eqnarray*}  
  with the densities $\hat\rho_f$ and $\hat\rho_s$, 
the viscosity $\nu_f$, the Lam\'e parameters 
$\mu_s$, $\lambda_s$ and the deformation gradient $\hat F$, and its
determinant $\hat J$. The stress tensors for the fluid and structure are
implemented by 
\[
\hat\sigma_f = -\hat pI + \hat\rho_f\nu_f (\hat\nabla\hat v \hat F^{-1}
  + \hat F^{-T} \hat\nabla\hat v^T)
\], 
and 
\[
\hat\sigma_s = \hat F (\lambda_s \text{tr}\hat E I + 2\mu_s \hat E)
\]
with $\hat E = \frac{1}{2}(\hat F^T \hat F - I)$.
\end{Problem}
The ALE approach belongs to interface-tracking methods in which 
the mesh is moved such that it fits in all time steps with 
the FSI-interface. However, this leads to 
a degeneration of the ALE map. Methods to circumvent such as 
degeneration as long as possible are re-meshing techniques or 
to use a biharmonic mesh motion technique.

@@

With the ALE code implemented in Example 
\ref{PDE_Stat_FSI_STVK} it is possible to treat fluid 
problems as well as FSI computations. 
In the case of fluid problems the deformation
gradient and its determinant become:
\begin{equation*}
F:= I , \quad \det F = J = 1.
\end{equation*}

@@

The quantities of interest are evaluations of 
$x$- and $y$ displacement at the point $A(0) = (0.6,0.2)$
and the drag and lift forces acting on the cylinder and the elastic beam:
\begin{align}
\label{drag_lift_forces}
(F_D , F_L) 
= {\int_{S_f} \sigma_f \cdot n_f \, ds + 
\int_{\Gamma_i} \sigma_s \cdot n_s \, ds},
\end{align}
where $S_{f}$ denotes the path over the cylinder in the fluid part and
$\Gamma_i$ the interface between the elastic beam and the 
fluid.

% \begin{table}[h]
%   \small
%   \centering
%     \begin{tabular}{llllcccc}    
%       \hline
%       Configuration & Date         & Refinement & Time dis. & $k$ & $\Delta p$& $F_D$ & $F_L$   \\ \hline\hline
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & BE        & 1.0    & 0.117412  &       &         \\
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & FS        & 1.0    & 0.117412  &       &         \\
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 0.117412  &       &         \\ 
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN        & 1.0    & 0.117383  &       &         \\ 
%       BFAC 2D-2     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 2.31541   &       &         \\ 
%       BFAC 2D-2     & Mar 10, 2010 & 3 global   & CN(k)     & 1.0    & 2.32011   &       &         \\   
%       BFAC 2D-2     & Mar 12, 2010 & 3 global   & CN(k)     & 1.0e-2 & 2.50288   &       &         \\   
%     \end{tabular}
%  \end{table}
% 
%\subsubsection{Code validation for FSI with ALE}
%
%The results are summarized below:
%
% \begin{table}[h]
%   \small
%   \centering
%     \begin{tabular}{llllccccc}    
%       \hline
%       Config. & Date & Refinement & Time dis. & $k$ &  $u_x(A) [\times 10^{-5}]$ &$u_y(A) [\times 10^{-4}]$ & $F_D$ & $F_L$  \\ \hline
%       FSI 1   & Mar 12, 2010 & 3 global & BE   & 1.0 & 8.2003  & 2.2732  & &        \\
%       FSI 1   & Apr 08, 2010 & 2 global & BE   & 1.0 & 8.2258  & 2.2813  & &        \\
%       FSI 1   & Apr 08, 2010 & 2 global & CN(k)& 0.5 & 8.2268  & 2.2813  & &        \\
%       FSI 1   & Apr 08, 2010 & 2 global & CN   & 0.5 & 8.2268  & 2.2813  & &        \\
%     \end{tabular}
%  \end{table}
%
%\vspace{0.2cm}
%
\textbf{Program description}

>>>>

In the present example, we solve a 
nonstationary fluid-structure interaction problem in 
arbitrary Lagrangian-Eulerian (ALE) coordinates. 
The mesh motion model is based on solving a biharmonic equation
\cite{Wi11} rather than a linear-elastic model. 
The underlying equations are stated in the following:
\begin{Problem}[FSI with biharmonic mesh motion]
  \label{eq:fsi:ale:biharmonic}
  Find $\{\hat v,\hat u,\hat w,\hat p\} \in \{ \hat v^D + \hat {\cal V}^0\} 
\times \{ \hat u^D + \hat {\cal V}^0 \}\times \hat {\cal V} \times \hat {\cal
  L}$, such that $\hat v (0) = \hat v^0$ and $\hat u(0) = \hat u^0$, 
for almost all time steps $t$, and
  \begin{eqnarray*}
    \begin{aligned}
      (\hat J \hat\rho_f \partial_t \hat v,\hat\psi^v)_{\hat\Omega_f}  
      +(\hat\rho_f \hat J  (\hat F^{-1}(\hat
      v-\partial_t \hat u)\cdot\hat\nabla) \hat v),
      \hat\psi^v)_{\hat\Omega_f} &\\
      + (\hat J\hat\sigma_f\hat
      F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_f}
      - \langle \hat g, \hat\psi^v \rangle_{\hat\Gamma_N}&\\ 
      %
      + (\hat\rho_s \partial_t \hat v,\hat\psi^v)_{\hat\Omega_s}  
      + (\hat J\hat\sigma_s\hat F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_s}
%      - (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
%      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
      &= 0&&\forall\hat\psi^v\in \hat V^0,
      \\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      (\hat\alpha_u\hat w, \hat\psi^w)_{\hat\Omega_f}  + ( \hat\alpha_u\hat\nabla\hat
      u, \hat\nabla\hat\psi^w)_{\hat\Omega_f} +( \hat\alpha_u\hat\nabla\hat
      w, \hat\nabla\hat\psi^w)_{\hat\Omega_s}  
%-\langle \hat\alpha_u \hat n_f\hat \nabla \hat u,\hat\psi^w\rangle_{\hat \Gamma_i} 
      & = 0 &&\forall\hat\psi^w\in \hat V , \\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      \hat\rho_s (\partial_t\hat u-\hat v,\hat\psi^u)_{\hat\Omega_s}
      + (\hat\alpha_u \hat \nabla \hat w,\hat \nabla\hat\psi^u)_{\hat\Omega_f}
      %-\langle \hat\alpha_w \hat n_f\hat \nabla \hat w,\hat\psi^u\rangle_{\hat \Gamma_i}
      &=0&&\forall\hat\psi^u\in \hat V^0 ,\\    
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      (\widehat{\text{div}}\,(\hat J\hat F^{-1}
      \hat v_f),\hat\psi^p)_{\hat\Omega_f} 
      + (\hat p_s ,\hat \psi^p)_{\hat\Omega_s}
      &=0&&\forall\hat\psi^p\in \hat L,
    \end{aligned}
  \end{eqnarray*}  
  with the densities $\hat\rho_f$ and $\hat\rho_s$, 
the viscosity $\nu_f$, the Lam\'e parameters 
$\mu_s$, $\lambda_s$ and the deformation gradient $\hat F$, and its
determinant $\hat J$. The stress tensors for the fluid and structure are
implemented by 
\[
\hat\sigma_f = -\hat pI + \hat\rho_f\nu_f (\hat\nabla\hat v \hat F^{-1}
  + \hat F^{-T} \hat\nabla\hat v^T),
\] 
and 
\[
\hat\sigma_s = \hat F (\lambda_s \text{tr}\hat E I + 2\mu_s \hat E)
\]
with $\hat E = \frac{1}{2}(\hat F^T \hat F - I)$.
Finally, we notice that this problem is driven by a Dirichlet 
inflow condition. It is possible to add a gravity term $\hat f_f$ or 
$\hat f_s$, which 
would enter as a right hand side force 
\[
- (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
\]
into the problem.
\end{Problem}
The ALE approach belongs to interface-tracking methods in which 
the mesh is moved such that it fits in all time steps with 
the FSI-interface. However, this leads to 
a degeneration of the ALE map. Methods to circumvent such as 
degeneration as long as possible are re-meshing techniques or 
to use (as suggested here) a biharmonic mesh motion technique.

@@

With the ALE code implemented in Example 
\ref{PDE_Stat_FSI_STVK} it is possible to treat fluid 
problems as well as FSI computations. 
In the case of fluid problems the deformation
gradient and its determinant become:
\begin{equation*}
\hat F:= I , \quad \det \hat F = \hat J = 1.
\end{equation*}

@@

The quantities of interest are evaluations of 
$x$- and $y$ displacement at the point $A(0) = (0.6,0.2)$
and the drag and lift forces acting on the cylinder and the elastic beam:
\begin{align}
\label{drag_lift_forces}
(F_D , F_L) 
= {\int_{S_f} \hat\sigma_f \cdot \hat n_f \, ds + 
\int_{\hat\Gamma_i} \hat\sigma_s \cdot \hat n_s \, ds},
\end{align}
where $S_{f}$ denotes the path over the cylinder in the fluid part and
$\Gamma_i$ the interface between the elastic beam and the 
fluid.

% \begin{table}[h]
%   \small
%   \centering
%     \begin{tabular}{llllcccc}    
%       \hline
%       Configuration & Date         & Refinement & Time dis. & $k$ & $\Delta p$& $F_D$ & $F_L$   \\ \hline\hline
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & BE        & 1.0    & 0.117412  &       &         \\
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & FS        & 1.0    & 0.117412  &       &         \\
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 0.117412  &       &         \\ 
%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN        & 1.0    & 0.117383  &       &         \\ 
%       BFAC 2D-2     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 2.31541   &       &         \\ 
%       BFAC 2D-2     & Mar 10, 2010 & 3 global   & CN(k)     & 1.0    & 2.32011   &       &         \\   
%       BFAC 2D-2     & Mar 12, 2010 & 3 global   & CN(k)     & 1.0e-2 & 2.50288   &       &         \\   
%     \end{tabular}
%  \end{table}
% 
%\subsubsection{Code validation for FSI with ALE}
%
%The results are summarized below:
%
% \begin{table}[h]
%   \small
%   \centering
%     \begin{tabular}{llllccccc}    
%       \hline
%       Config. & Date & Refinement & Time dis. & $k$ &  $u_x(A) [\times 10^{-5}]$ &$u_y(A) [\times 10^{-4}]$ & $F_D$ & $F_L$  \\ \hline
%       FSI 1   & Mar 12, 2010 & 3 global & BE   & 1.0 & 8.2003  & 2.2732  & &        \\
%       FSI 1   & Apr 08, 2010 & 2 global & BE   & 1.0 & 8.2258  & 2.2813  & &        \\
%       FSI 1   & Apr 08, 2010 & 2 global & CN(k)& 0.5 & 8.2268  & 2.2813  & &        \\
%       FSI 1   & Apr 08, 2010 & 2 global & CN   & 0.5 & 8.2268  & 2.2813  & &        \\
%     \end{tabular}
%  \end{table}
%
%\vspace{0.2cm}
%
\subsubsection{Program description}

### 2013-12-13 Examples/PDE/StatPDE/Example10/content.tex
Update Doku

This example is an extension the previous one.
We solve an stationary FSI problem either with 
INH material (see Problem definition before) 
or St. Venant Kirchhoff material STVK: \index{structure} \index{structure!compressible St.Venant-Kirchhoff (STVK) material}
\begin{Problem}[Compressible Saint Venant-Kirchhoff, Lagrangian framework]
  \label{prob:stvk:lagrange}
  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
  \times \{ \hat u_f^D + \hat V\}$, 
  such that 
  \begin{eqnarray}
    \begin{aligned}   
      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
      &&\quad\forall\hat \phi^v\in\hat V_s\\
      %%%%%%%%%%%%%%%% 
      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
    \end{aligned}
  \end{eqnarray}
  where $\rho_s$ is the density of the structure, $\mu_s$ and $\lambda_s$ 
  the Lam\'e coefficients, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
  $\hat g_s^\sigma$ some forces on the interface.\index{Lam\'e coefficients} The properties of the STVK material
  is specified by the constitutive law
  \begin{eqnarray}
    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
  \end{eqnarray}
\end{Problem}
Often, the elasticity properties of structure materials is characterized by 
Poisson's ratio $\nu_s$ ($\nu_s < \frac{1}{2}$ \index{Poisson's ratio} for 
compressible materials) and the Young modulus $E$ \index{Young modulus}. The 
relationship to the Lam\'e coefficients $\mu_s$ and $\lambda_s$ is
given by:
\begin{eqnarray}
\nu_s = \frac{\lambda_s}{2(\lambda_s + \mu_s)}, \qquad 
E = \frac{\mu_s (\lambda_s + 2\mu_s)}{(\lambda_s + \mu_s )} . 
\end{eqnarray}
The whole equation system is solved on the benchmark configuration domain. For 
details on parameters and geometry, we refer to the 
numerical FSI benchmark \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!FSI benchmark} proposal from Hron and Turek [2006].

@@

Compared to the two former examples, there are some differences which we will briefly discuss in the following. First of all, the problem is nonlinear in contrast to the former ones. We work on a different domain (given in the \textit{benchfst0100tw.inp} file), namely a channel with a cylinder put at half height near the inflow boundary; further \textit{.inp} files yield the possibility to vary the domain.\\
Furthermore, in the \textit{dope.prm} parameter file there are two additional subsections which are added only for the current problem. From the denotation of these subsections one can immediately see where in the code the parameters are used.\\
As we want to compute certain benchmark quantities, we have to regard corresponding functionals in the \textit{functionals.h} file. The pressure at a point as well as the displacement in $x$- and $y$-directions are point values; furthermore we implement the drag and lift functionals (for which we need the additionally defined problem parameters).\\
As before, we build up the element and boundary equations and matrices in the \textit{localpde.h} file. Apart from using the additionally defined problem parameters and modeling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement. Further on, everything is pretty much the same as in the last example, but we have to use the \texttt{SetBoundaryFunctionalColors} function of the \texttt{PDEProblemContainer} class to be able to compute drag and lift in the respective functional classes in \textit{functionals.h}.\\
The main innovation in contrast to the preceding examples is the refinement of the grid combined with a simple error estimator given in the deal.II \texttt{KellyErrorEstimator} class. If we look at the output of our program, everything is computed several times (once on each refinement level). Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.

>>>>

This example is an extension the previous one.
We solve an stationary FSI problem either with 
INH material (see Problem definition before) 
or St. Venant Kirchhoff material STVK: \index{structure} \index{structure!compressible St.Venant-Kirchhoff (STVK) material}
\begin{Problem}[Compressible Saint Venant-Kirchhoff, Lagrangian framework]
  \label{prob:stvk:lagrange}
  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
  \times \{ \hat u_f^D + \hat V\}$, 
  such that 
  \begin{eqnarray}
    \begin{aligned}   
      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
      &&\quad\forall\hat \phi^v\in\hat V_s\\
      %%%%%%%%%%%%%%%% 
      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
    \end{aligned}
  \end{eqnarray}
  where $\rho_s$ is the density of the structure, $\mu_s$ and $\lambda_s$ 
  the Lam\'e coefficients, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
  $\hat g_s^\sigma$ some forces on the interface.\index{Lam\'e coefficients} The properties of the STVK material
  is specified by the constitutive law
  \begin{eqnarray}
    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
  \end{eqnarray}
\end{Problem}
Often, the elasticity properties of structure materials is characterized by 
Poisson's ratio $\nu_s$ ($\nu_s < \frac{1}{2}$ \index{Poisson's ratio} for 
compressible materials) and the Young modulus $E$ \index{Young modulus}. The 
relationship to the Lam\'e coefficients $\mu_s$ and $\lambda_s$ is
given by:
\begin{eqnarray}
\nu_s = \frac{\lambda_s}{2(\lambda_s + \mu_s)}, \qquad 
E = \frac{\mu_s (\lambda_s + 2\mu_s)}{(\lambda_s + \mu_s )} . 
\end{eqnarray}
On fluid side, we extend the problem from Stokes flow 
to stationary Navier-Stokes flow to demonstrate how to 
implement the nonlinear convection term
\[
             (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
             \hat\phi^v)_{\hat\Omega_f}.
\]

The whole equation system is solved on the benchmark configuration domain. For 
details on parameters and geometry, we refer to the 
numerical FSI benchmark \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!FSI benchmark} proposal from Hron and Turek [2006].

@@

Compared to the previous Example \ref{PDE_Stat_FSI_STVK}, 
there are some differences which we will briefly discuss in the following. First of all, the problem is nonlinear in contrast to the former ones. We work on a different domain (given in the \textit{benchfst0100tw.inp} file), namely a channel with a cylinder put at half height near the inflow boundary; further \textit{.inp} files yield the possibility to vary the domain.\\
Furthermore, in the \textit{dope.prm} parameter file there are two additional subsections which are added only for the current problem. From the denotation of these subsections one can immediately see where in the code the parameters are used.\\
As we want to compute certain benchmark quantities, we have to regard corresponding functionals in the \textit{functionals.h} file. The pressure at a point as well as the displacement in $x$- and $y$-directions are point values; furthermore we implement the drag and lift functionals (for which we need the additionally defined problem parameters).\\
As before, we build up the element and boundary equations and matrices in the \textit{localpde.h} file. Apart from using the additionally defined problem parameters and modeling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement. Further on, everything is pretty much the same as in the last example, but we have to use the \texttt{SetBoundaryFunctionalColors} function of the \texttt{PDEProblemContainer} class to be able to compute drag and lift in the respective functional classes in \textit{functionals.h}.\\
The main innovation in contrast to the preceding examples is the refinement of the grid combined with a simple error estimator given in the deal.II \texttt{KellyErrorEstimator} class. If we look at the output of our program, everything is computed several times (once on each refinement level). Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.

### 2013-12-13 Examples/PDE/StatPDE/Example4/content.tex
Update Doku

>>>>

\begin{remark}[Why this example?]
Originally, \dope{} was designed for coupled and nonlinear problems
with possible PDE-based optimization extensions. Later, we decided 
to add the most simplest PDE (the Laplace/Poisson equation) 
in order to demonstrate how DOpE treats
this well-known example. In addition, the first-time user might 
start here to get a feeling for \dope{} and its capabilities.
\end{remark}
